---
title: "DC Crime"
output: html_document
author:
- Tian Yang
- Huixin Rao
- Hanbo Li
---

```{r , echo=FALSE}
library(dplyr)
library(class)
library(caret)
library(data.table)
library(rpart)
library(randomForest)
```

Combine the raw data together.
```{r}
data08<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2008.csv")
data09<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2009.csv")
data10<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2010.csv")
data11<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2011.csv")
data12<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2012.csv")
data13<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2013.csv")
data14<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2014.csv")
data15<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2015.csv")
data16<-read.csv("/Users/yangtian/Desktop/ds final/Crime_Incidents__2016.csv")
data <- rbind(data08, data09, data10, data11,data12,data13,data14,data15,data16)
dim(data)
head(data)
```

## Data Preprocessing

### The transfer of date 
```{r}
#transfer the date variable into year, month, day, hour and day_of_week, which are easier for us to analyze
week<-weekdays(as.Date(data$START_DATE))

#there are a lot of na in day_of_week, we need to transfer them into usable value
table(is.na(week))
sd <- substring(data$START_DATE,1,10)
data <- data[-which(sd == ""),] #delete 13 na values (13 is a small number in our case)
sd<- substring(data$START_DATE,1,10)
week <- weekdays(as.Date(data$START_DATE))
table(is.na(week))
wrong<-sd[is.na(week)]
head(wrong)
nwd<-weekdays(as.Date(wrong))
week[is.na(week)]<-nwd
st_y<-substring(data$START_DATE, 1,4)
st_m<-substring(data$START_DATE, 6,7)
st_d<-substring(data$START_DATE, 9,10)
st_h<-substring(data$START_DATE, 12,13)

#add some new variables in the dc_crime dataset
data$start_year<-st_y
data$start_month<-st_m
data$start_hour<-st_h
data$start_dayOfWeek<-week
data$START_DATE<-NULL

#data$season[data$start_month>=3 & data$start_month<=5]<-'Spring'
#data$season[data$start_month>=6 & data$start_month<=8]<-'Summer'
#data$season[data$start_month>=9 & data$start_month<=11]<-'Autumn'
#data$season[data$start_month==1|data$start_month==2|data$start_month==12]<-'Winter'
dim(data)
```

### Delete Overlaping Variables 
Since there are too many varables describing location, we choose NEIBORHOODCLUSTER as our location varable. 
```{r }
colnames(data)
data$CCN<-NULL
data$REPORTDATETIME<-NULL
data$LASTMODIFIEDDATE<-NULL
data$BLOCKSITEADDRESS<-NULL
data$CENSUS_TRACT<-NULL
data$END_DATE<-NULL
data$VOTING_PRECINCT<-NULL
data$CENSUS_TRACT<-NULL
data$PSA<-NULL
data$ANC<-NULL
data$WARD<-NULL
data$BUSINESSIMPROVEMENTDISTRICT<-NULL
data$DISTRICT<-NULL
data$BLOCK_GROUP<-NULL
data$OBJECTID<-NULL
data$X<-NULL
data$Y<-NULL
summary(data)
```

## Model Selection

### KNN

#### Step 1 : Transformming all data type to numeric

```{r}
shift<-as.character(data$SHIFT)
shift[which(shift == "DAY")]<-1
shift[which(shift == "EVENING")]<-2
shift[which(shift == "MIDNIGHT")]<-3
table(shift)
offense<-as.character(data$OFFENSE)
offense[which(offense == "ARSON")]<-1
offense[which(offense == "ASSAULT W/DANGEROUS WEAPON")]<-2
offense[which(offense == "BURGLARY")]<-3
offense[which(offense == "HOMICIDE")]<-4
offense[which(offense == "MOTOR VEHICLE THEFT")]<-5
offense[which(offense == "ROBBERY")]<-6
offense[which(offense == "SEX ABUSE")]<-7
offense[which(offense == "THEFT F/AUTO")]<-8
offense[which(offense == "THEFT/OTHER")]<-9
table(offense)
method<-as.character(data$METHOD)
method[which(method == "GUN")]<-1
method[which(method == "KNIFE")]<-2
method[which(method == "OTHERS")]<-3
table(method)
dw<-as.character(data$start_dayOfWeek)
dw[which(dw == "Monday")]<-1
dw[which(dw == "Tuesday")]<-2
dw[which(dw == "Wednesday")]<-3
dw[which(dw == "Thursday")]<-4
dw[which(dw == "Friday")]<-5
dw[which(dw == "Saturday")]<-6
dw[which(dw == "Sunday")]<-7
table(dw)

data_knn <- data
data_knn$SHIFT<-as.numeric(shift)
data_knn$start_dayOfWeek<-as.numeric(dw)
data_knn$METHOD<-as.numeric(method)
data_knn$OFFENSE<-as.numeric(offense)
summary(data_knn)
```

#### Step 2 : Set train and test data
```{r}
set.seed(1)
test <- sample_frac(na.omit(data_knn),0.3)
train <- sample_frac(na.omit(data_knn),0.7)
test.y<-test[,2]
train.y<-train[,2]
test.x<-test[,-2]
train.x<-train[,-2]
```


#### Step 3 : Do the KNN prediction

```{r}
knn.pred=knn(train.x,test.x,train.y,k=1)
table(knn.pred,test.y)
confusionMatrix(knn.pred,test.y)
```
 
We could see that the accuracy of KNN is very high, which is over 0.8.

### Decision Tree and Random Forest

```{r}
crime <- data
crime <- crime[-c(which(data$start_year == "1993"), which(data$start_year == "1997"), which(data$start_year == "1800")),]
crime$start_year<-as.integer(crime$start_year)
crime$start_month<-as.integer(crime$start_month)
crime$start_hour<-as.integer(crime$start_hour)
crime$start_dayOfWeek<-as.factor(crime$start_dayOfWeek)

# set the train and test data the same as KNN
set.seed(1)
dt = sort(sample(nrow(crime), nrow(crime)*.7))
train<-crime[dt,]
train <- train[!is.na(train$NEIGHBORHOODCLUSTER),]
test<-crime[-dt,]

#DECISITON TREE
treemodel <- rpart(OFFENSE ~METHOD+SHIFT+start_year+start_month+start_hour+start_dayOfWeek
                   +NEIGHBORHOODCLUSTER,   data=train)
printcp(treemodel)

# plot the tree
plot(treemodel, uniform=TRUE, 
     main="Classification Tree ")
text(treemodel, use.n=TRUE, all=TRUE, cex=.8)

# confusion matrix
Prediction <- predict(treemodel, test, type = "class")
confusionMatrix(Prediction, test$OFFENSE)

# random forest
fitrf <- randomForest(OFFENSE ~SHIFT+METHOD+start_year+
                        start_month+start_hour+start_dayOfWeek+
                        NEIGHBORHOODCLUSTER,   data = train, ntree=100)
summary(train)
print(fitrf) # view results 
importance(fitrf) # importance of each predictor
varImpPlot(fitrf)
test$pred_crime<-predict(fitrf,test)
confusionMatrix(test$pred_crime, test$OFFENSE)
```

The accuracy of Decision tree and Random Forest are both around 0.5, which is not bad, but comparing to KNN, it's worse.

## Predict the Crime Probability 

We need to predict the crime probability in a certain location within a certain period of time, in order to make our results meaningful.

```{r}
data$start_dayOfWeek <- as.character(data$start_dayOfWeek)

# this function needs three inputs : date, period of a day and location, and it will give a crime probability output.
pred <- function(date, shift, location){
  ss<-gsub(" ","",deparse(substitute(shift)))
  ll<-as.integer(deparse(substitute(location)))
  dd<-gsub(" ","",deparse(substitute(date)))
  shift1 <- data[which(data$SHIFT == ss),]
  location1 <- shift1[which(shift1$NEIGHBORHOODCLUSTER == ll),]
  dow <- weekdays(as.Date(dd))
  r<-dim(data[which(data$start_dayOfWeek == dow),])[1]/dim(data)[1]
  p <- dim(location1)[1]/dim(data)[1]*r
  t<-c(substring(dd,1,4),substring(dd,6,7),substring(dd,9,10),ss,ll, dow, p)
  return(t)
}
```

## Display the Results

We will show the predict results around GWU, which is in neighborhood 5.
```{r}
# predict the crime type around GWU, which is in neighborhood 5
t1<-pred(2016-12-17, DAY, 05)
t2<-pred(2016-12-17, NIGHT, 05)
t3<-pred(2016-12-17, MIDNIGHT, 05)
t4<-pred(2016-12-18, DAY, 05)
t5<-pred(2016-12-18, NIGHT, 05)
t6<-pred(2016-12-18, MIDNIGHT, 05)
t7<-pred(2016-12-19, DAY, 05)
t8<-pred(2016-12-19, NIGHT, 05)
t9<-pred(2016-12-19, MIDNIGHT, 05)
t<-as.data.frame(rbind(t1,t2,t3,t4,t5,t6,t7,t8,t9))
colnames(t)<-c("year","month","day","shift","neighborhood", "day of week", "crime probability")

# build up new test data, which we are going to predict
colnames(test.x)
SHIFT<-c(1,2,3,1,2,3,1,2,3)
METHOD <- c(1,1,1,2,2,2,3,3,3)
BLOCKXCOORD <- rep(2374578,9)
BLOCKYCOORD <- rep(820740,9)
NEIGHBORHOODCLUSTER <- rep(05,9)
start_year <- rep(2016,9)
start_month <- rep(12,9)
start_dayOfWeek <- c(6,6,6,7,7,7,1,1,1)
start_hour <- c(10,20,04,08,22,02,15,23,01)

newtest <- as.data.frame(cbind(SHIFT, METHOD,BLOCKXCOORD,BLOCKYCOORD,NEIGHBORHOODCLUSTER,start_year,start_month,start_hour,start_dayOfWeek))

# use knn to predict the crime type
knn.pred=knn(train.x,newtest,train.y,k=1)
crime_type <- knn.pred

# use the probability function to predict the crime probability
probability <- t$`crime probability`
results <- cbind(t,crime_type, probability)
results

```

According to this result, for example, there is a 0.1% probability that there will be a type 3 crime (BURGLARY) happen in day time in GWU neighborhood.
